Marine species do not need visas: the transboundary nature of the world’s exploited marine species

### 2020-03-06 Manuscript_Draft.Rmd
Changes in the manuscript for submission

---
title: "Marine species do not need visas: the transboundary nature of the world’s exploited marine species"
author: "Authors: Juliano Palacios-Abrantes^1^, Gabriel Reygondeau^1,2^, Colette Wabnitz^1,3^ and William W.L. Cheung^1^"
date: "Affiliations: ^1^Institute for the Oceans and Fisheries, University of British Columbia, Vancouver, Canada. ^2^Yale University, New Haven, Connecticut, USA. ^3^Stockholm Resilience Center, Stockholm University, Stockholm, Sweden."
# subtitle: "[DRAFT DO NOT SHARE]"
subtitle: "Most neighbor nations share marine exploited species"
output:
  word_document:
    reference_docx: ./Manuscript_Instructions/science_template.docx
  pdf_document:
    fig_caption: yes
geometry: margin=1in
editor_options:
  chunk_output_type: inline
bibliography: ./References/FishForVisa_Ref.bib
csl: ./References/science.csl
---

@@

<!-- - Journal: *Science* -->
<!-- - Information to authors (https://www.sciencemag.org/authors/science-information-authors) -->
<!-- - There are two types of article: -->

<!-- -- **Reports** (up to ~2500 words including references, notes and captions–corresponds to ~3 printed pages in the journal) present important new research results of broad significance. Reports should include an abstract, an introductory paragraph, up to four figures or tables, and about 30 references. -->

<!-- Materials and Methods should be included in supplementary material -->

<!-- --- Instructions for preparing an initial manuscript https://www.sciencemag.org/authors/instructions-preparing-initial-manuscript -->

@@

<!-- - **Potential Reviewers (J)**   -->
<!--  -- Gakushi Ishimura, Faculty of Agriculture Department of Food Production and Environmental Management, Iwate University, Morioka, Japan; gakugaku@iwate-u.ac.jp -->
<!--  -- Rebecca Selden, Wellesley College, Wellesley, MA, USA; bselden@wellesley.edu -->
<!--  -- Julia Xue, Shanghai Jiao Tong University; juliaxue@ouc.edu.cn -->
<!--  -- Claire Armstrong, The Norwegian College of Fishery Science, Tromso, Norway claire.armstrong@uit.no -->

@@

# **One Sentence Summary**
<!-- *A brief teaser statement highlighting main result of the paper, understandable by a scientist not in your field, without jargon or abbreviations. This will appear online adjacent to the title and should not repeat phrases already present there* -->

Marine species shared by neighboring nations contribute to over 50% of catches and fishing revenue with direct implications for fisheries management.

<!-- # Field code (Max. 2) -->

<!-- - *Fisheries* -->

# **Abstract**
<!-- *An opening sentence that sets the question that you address and is comprehensible to the general reader, background content specific to this study, results, and a concluding sentence. It should be a single paragraph.* -->

Regulatory boundaries and species distributions are often not aligned, especially in the case of marine species crossing multiple Economic Exclusive Zones (EEZs). This represents a challenge for fisheries management as policies tend to focus at the national level, yet collaborations are needed to maximize long-term ecological, social and economic benefits. Here, we combined species distributions and the spatial delimitation of EEZs to identify the number of commercially exploited marine species that are shared between neighboring nations. We found that 67% of the world's identified fished species are transboundary (n = 633) and were responsible for 80% of the global-EEZs catches (48 million tons) and 78.5% of fisheries revenue (USD 65 billion) in the last decade. Some countries received over 90% of their catch and economic benefits from a few shared resources. This study highlights the large and under-recognized challenge in achieving international fisheries management goals, particularly as species redistribute under climate change.

# **Main Article**
<!-- # Introduction (~555/500 words) -->

@@

We overlaid the known distribution of 938 commercially valuable marine species responsible for an average of 96.5% of global EEZ catch reconstructed by the *Sea Around Us* at the species level between 2005 and 2014, and 241 EEZs of 198 coastal territories. We define a ‘stock’ unit as a species in an EEZ, instead of using a population ecology definition (genetically or morphological distinct units) due to lack of such biological information for almost all the species included in this study [@Teh:2015gd]. We rely on multiple data sources including occurrence, distribution models and catch data and only considered a species to be present in a grid cell if all data sources predicted positive occurrence (see Methods).

<!-- ## Results and discussion (~1000 words) -->

@@

![Fig. 1. Number of transboundary species and their contribution to global fisheries catches and revenue. A) The number of species and amount of revenue are represented by color coding of EEZs and land polygons, respectively. B) Contribution of transboundary species to regional revenue (left) and catch (right). Regions according to the United Nations sub-regions. Points = mean +- sd. Revenue in 2010 real USD.](./Figures/Fig1.png)

We group countries according to the United Nations sub-regions and found that transboundary species are particularly economically important for Northern America (mean revenue = 3,901 $\pm$ 4,497 million USD) and Eastern Asia (2,827 $\pm$ 2,667 million USD). The estimated revenues from transboundary species in these two regions are significantly different from other regions in terms of revenue (one-way ANOVA; *DF*(16,158) = 4.948, *p* < 0.001, $\alpha$ = 0.05) (Fig 1B). China (USD 6,279 million) in E. Asia and the USA (USD 9,605 million) contribute 55% and 82% to the total revenues from the E. Asia and N. America regions, respectively. Russia (USD 6,361 million), Peru (USD 5,211) and Japan (USD 3,368) are amongst the top five countries with the most fishing revenue generated from transboundary species in the last decade. These five nations are responsible for 40% of the global fisheries revenue from transboundary species. Peru and Russia, responsible for the two largest fisheries in the world [@FAO:2018KK], lead transboundary catch statistics harvesting over 5.8 million tons annually each on average between 2005 and 2014 (Fig S1). Peru's catches are mainly composed of Peruvian anchoveta (*Engraulis ringens*), shared with Chile, and account for 79% of the transboundary species production. Similarly, Russia shares the Alaskan pollock (*Theragra chalcogramma*) fishery with both Japan and the USA. Transboundary species also contribute largely to fisheries in E. Asia (one-way ANOVA; *Df*(16,158) = 2.265, *p* = 0.005, $\alpha$ = 0.05). Particularly, China the world's top fish producer [@FAO:2018KK] getting one third (5.1 million tons) of the total catches from transboundary species, followed by Japan (1.8 million tons) and South Korea (1.06 million tons). However, when we account for the total area of waters in different EEZs and regions relative to their catches, the importance of transboundary species to European fisheries become apparent (Fig. S2, Table S1).

We determined the exploitation category of each species according to the catch trend within each EEZ (Fig. 2). Previous studies demonstrated that catch trends can be used to infer whether a stock is healthy, over-exploited or collapsed [@Kleisner:2011wn]. We found significant differences in exploitation categories based on whether or not a species was transboundary (MANOVA, *Df*(2,447) = 27.091, *p* < 0.001). Specifically, transboundary species that are only fished by one country are generally less likely to be over-exploited compared to those shared by neighboring countries (Table S2). Significant differences were found for all categories within transboundary species (one-way ANOVA, *Df*(2,447) = 55.01, *p* $\leq$ 0.001, $\alpha$ = 0.05; Tukey's post hoc test *p* < 0.05) but only Category C for discrete species (one-way ANOVA, *Df*(2,380) = 21.86, *p* < 0.001; $\alpha$ = 0.05; Tukey's post hoc test for Categories C-A and C-B *p* < 0.001, and B-A 0.052). This result suggests that the management of discrete species could be better off as the variation between Categories A and B is closer (*p* = 0.052) than for transboundary species (*p* = 0.001).

@@

![Fig. 2. Number of transboundary species by catch category. Category A, fishing between 10%-50% of maximum catch level after a drop larger than 10% of the recorded maximum catch; Category B, fishing over 50% of maximum catch level; Category C, fishing less than 50% of maximum catch](./Figures/Fig2.png)

Empirical analysis suggests that cooperation in the management of transboundary stocks will have better outcomes, although there are specific cases where cooperation might not be the best strategy [@Munro:2002uf]. Examples of successful cooperative-management include Norway and Russia over Atlantic cod (*Gadus morhua*) [@Eide:2013dha] and Namibia and South Africa over hake (*Merluccius spp*) [@Sumaila:2003vw]. Lack of collaboration over shared stocks may threaten stock sustainability, reduce the future profitability potential of the fishery, and create conflict between neighboring nations [@Clark:1980js; @Spijkers:2017ij]. Transboundary fisheries are abundant and of economic importance for some of the top countries with past fisheries-related conflicts including Canada, the USA, Russia and the European Union (EU) [@Spijkers:2019cz]. For example, since 2007, the EU, Norway, Iceland, and the Faroe Islands have been at odds over the size and relative allocation of total allowable catches for Atlantic mackerel (*Scomber scombrus*) due to a climate-driven distribution shift [@Spijkers:2017ij]. Mackerel is a transboundary species that contributed an average of 598.4 thousand tons to catches (8.19% of total) and USD 850.2 million in revenue (7.17% of total) to the countries involved in the dispute between 2005 and 2014. The relatively small contribution of mackerel to the involved countries suggests that even a small fraction of the year's catch can lead to important conflicts between neighboring countries. Climate change is expected to continue changing the distribution and share proportion of fish stocks among countries up to a degree that new transboundary species will emerge [@Pinsky:2018cb], while some species will exit EEZs [@Oremus:2020id]. Exploring the detailed impacts of climate change in share species is key to develop local adaptation methods to minimize the effect of such distributional change on the conservation of species as well as the economy and food security of coastal nations.

Over 130 marine fish species occur in tropical and subtropical waters around the world [@Nelson:2016kk], from highly migratory species associated to pelagic-oceanic ecosystems like tunas (*Thunnus sp.*), to less mobile reef-associated species like greater amberjack (*Seriola dumerili*), to species found in demersal ecosystems like gilthead seabream (*Sparus aurata*). Species associated to pelagic-oceanic ecosystems are the only group whose EEZ range (i.e., the number of EEZs where the species occur as transboundary) was significantly different from other groups (one-way ANOVA; *DF*(5,597) = 53.82, *p* < 0.001, $\alpha$ = 0.05; Tukey's post hoc test *p* < 0.05) with a median of 40 EEZs per species. The median for species of all other ecosystem preferences is close to, or less than, 20, as many of these species have a narrower distribution or are less mobile (Fig. 3).

@@

img <- readPNG(here("./Figures/Fig3.png"))
grid.raster(img)
```

![Fig. 3. Number of EEZs shared by transboundary species. A) Number of EEZs by species organized by associated habitat according to FishBase. Showing only species that share > 20 EEZs. B) Catchs of main fishing countries fishing each ecosystem preference's top five species in terms of shared EEZs (shown in bold in A) between 2005 and 2014.](./Figures/Fig3.png)

Due to their large migration patterns and presence in areas beyond national jurisdiction, many highly migratory species are managed by Regional Fisheries Management Organizations (RFMOs). Fifteen major RFMOs in the world oversee the management of 99 species [@CullisSuzuki:2010fi]. We estimated that 81.8% of these species are shared across neighboring RFMOs, with the International Commission for the Conservation of Atlantic Tunas (ICCAT) sharing over 40 species with neighboring RFMOs (Fig. S3). The effectiveness of RFMOs has been previously questioned [@CullisSuzuki:2010fi]. However, despite a global expansion of the industrial fleet in the last decade [@Coulter:2020gr], a recent analysis of 40 tuna and billfish stocks suggests that most stocks are sustainable harvested with only three currently considered as overfished [@Pons:2016ep]. In the other hand, species with narrower ranges are often managed at the national level due to localized populations [@Payne:2008vm]. However, the potential boundary-crossing of these species must not be ignored. The world of fisheries is highly connected through larvae dispersal and so, even centralized populations could be supplying populations thousands of kilometers away [@Ramesh:2019va]. In addition, coral regions often share multiple species, making coordination among countries key to achieve fisheries and conservation goals [@Payne:2008vm]. Many coastal communities, particularly in developing regions, are highly dependent on subsistence fishing for food security and livelihoods [@CisnerosMontemayor:2016gq; @Wabnitz:2018gf]. For example, all of the top five countries fishing transboundary reef and pelagic-oceanic associated species (Fig. 3B) are among the highest fish consumption countries [@FAO:2018KK]. In these countries, both pelagic and reef fishes contribute substantially to both local food security and livelihood [@Cabral:2018hc]. In the case of Kiribati, the Solomon Islands and Papua New Guinea, they are parties to the Pacific Nauru Agreement that sustainable manage 30% of global skipjack tuna [@Aqorau:2018bh], estimated to produce $2 billion worth of fish every year, representing a role model for transboundary fisheries management [@Hanich:2010ht].

@@

# **References**

# **Acknowledgments**

We thank T. Clark and A. Tidd for helpful comments and discussions. **Funding**: G.R. was supported by the EO Wilson foundation. C.W. was supported by the Walton Family Foundation (grant 2018-1371), the David and Lucile Packard Foundation (grant 2019-68336), and the Gordon and Betty Moore Foundation (grant GBMF5668.02). W.W.L.C was supported by the NSERC Discovery Grant and the SSHRC Partnership. **Authors contributions**: J.P.A. Conceptualization, methodology, formal analysis, writing original draft. G.R. Methodology, data curation, writing - review & editing. C.W. Methodology, writing - review & editing. W.W.L.C. Supervision, writing - review & editing. All authors provided valuable input to the study conceptualization and discussed the results and analysis. **Competing interests**: No declared. **Data and materials availability**: The Sea Around Us datasets used are publicly available at http://www.seaaroundus.org/, the occurrence dataset is all publicly available (see Methods), the ENM-Nereus data can be accessed upon request to G.R. All code is available at https://github.com/jepa/FishForVisa

## Data and materials availability
All data is publicly available online and upon request from the authors.

## List of Supplementary Materials
Figures S1-S4
Tables S1 -S2
References in main manuscript (1-27)

# **Materials and Methods**

@@

We developed a four-criteria methodology to determine whether or not a species can be considered transboundary. Only species that meet all criteria were considered as transboundary and the analysis was undertaken only within the EEZs of coastal states. In some cases, criteria can also be interpreted as indices to measure the uncertainty in the analysis as they provide a continuous index (e.g. from 0 to 1) rather than a discrete, yes or no, index. All analyses were run using the statistical software *R version 3.5.2 (2018-12-20)* with the packages *data.table* [@Packagedatatable:2019uh], *janitor* [@PackagejanitorSim:2018ur], *wesanderson* [@Packagewesanderson:2018vn], *rfishbase* [@PackagerfishbaseR:2019th], *R.matlab* [@PackageRmatlabRe:2018un], *sf* [@PackagesfSimpleF:2018vp], *sp* [@PackagespClasses:2019tn], *tidiverse* [@PackagetidyverseE:2017vq], *tidytext* [@PackagetidytextTe:2019vk], and *zoo* [@PackagezooSInfr:2019tm]. All code is available at https://github.com/jepa/FishForVisa.

@@

We estimated the economic contribution in 2010 real USD of transboundary species for each country using global ex-vessel price data [@Tai2017]. The database we draw from includes ex-vessel price from multiple sources and a structured interpolation method (e.g. similar countries, species) to fill in data gaps [@Sumaila:2015uc]. The contribution of transboundary species was derived by aggregating the catch of species occurring within the EEZs of countries sharing that species. For this study, we did not include catch nor revenue from areas beyond national jurisdiction. Finally, we used the catch data to determine the exploitation category of each species within an EEZ. Although this method has previously been used to estimate stock status [@Grainger:1996tk], the categories presented here are intended as catch trends, and not fishing status for each species as many environmental and social-economic factors (e.g. temperature, markets, fishing policies) affect fishing effort and catch volume [@Pauly:2013gs; @Branch:2008kk]. We only assessed species within each EEZ for which at least ten years of data were available between the first and last reported catch and with at least five consecutive years of data. Three final categories were drawn for each species depending on the catch volume within each EEZ (e.g. present, maximum and minimum EEZ's historical catch) and the year (e.g. year of maximum historical catch of the species within that EEZ) (Table 2) [@Kleisner:2011wn]. Following this concept, Category A would represent fisheries that are re-building or under-exploited, Category B would be species that are sustainable harvested, and Category C would be species that are over-exploited. Finally, we reported the predominant category over the last ten years of catch data (2005-2014). We performed a series of one-way analysis of variance (one-way ANOVA) and Multivariate analysis of variance (MANOVA) to determine statistically significant differences between the means of different groups (e.g., geographical regions, species, exploitation categories) of transboundary species and their contribution to catch and fishing revenue [@Chambers:1992up; @Krzanowski:1990wm]. We used the standard *R* functions for the ANOVA and MANOVA after testing for assumptions.

@@

@@

# **Supplements**

@@

![Number of transboundary species per EEZ and their contributions to countries’ EEZ catch. The number of species is displayed in the EEZ while catch (in thousand tons) is displayed in the land polygon.](./Figures/FigS1.png)

Differences in transboundary fisheries can also be illustrated in terms of catch-revenue/EEZ $km^2$ (Figure S2). As an example, the aggregated area of all EEZs of Northern Europe is 3.3 million $Km^2$, the $5^{th}$ smallest of the 17 groups analyzed (Table S1), however, it has the most revenue (USD 246.7 million per $Km^2$) and catch (260.1 tons per $Km^2$) of transboundary species per EEZ area, but only significantly differing from Sub-Saharan African and Southern Europe countries (*p* < 0.05).

>>>>

---
title: "Marine species do not need visas: the transboundary nature of the world’s exploited marine species"
author: "Authors: Juliano Palacios-Abrantes*^1^, Gabriel Reygondeau^1,2^, Colette Wabnitz^1,3^ and William W.L. Cheung^1^"
date: "Affiliations: ^1^Institute for the Oceans and Fisheries, University of British Columbia, Vancouver, Canada. ^2^Yale University, New Haven, Connecticut, USA. ^3^Stockholm Resilience Center, Stockholm University, Stockholm, Sweden."
# subtitle: "[DRAFT DO NOT SHARE]"
subtitle: "Most neighbor nations share marine exploited species"
output:
  word_document:
    # reference_docx: ./Manuscript_Instructions/science_template.docx
  pdf_document:
    fig_caption: yes
geometry: margin=1in
editor_options:
  chunk_output_type: inline
bibliography: ./References/FishForVisa_Ref.bib
csl: ./References/nature_cc.csl
---

@@

<!-- - Journal: *Scientific Reports* -->
<!-- - Information to authors (https://www.nature.com/srep/author-instructions/submission-guidelines) -->
<!-- - There are two types of article: -->

<!-- -- Articles should be no more than 11 typeset pages in length. As a guide, the main text (not including Abstract, Methods, References and figure legends) should be no more than 4,500 words. The maximum title length is 20 words. The Abstract (without heading) - which must be no more than 200 words long and contain no references - should serve both as a general introduction to the topic and as a brief, non-technical summary of the main results and their implications.. -->

@@

<!-- - **Potential Reviewers (J)**   -->
<!--  -- Claire Armstrong, The Norwegian College of Fishery Science, Tromso, Norway claire.armstrong@uit.no -->
<!--  -- Gakushi Ishimura, Faculty of Agriculture Department of Food Production and Environmental Management, Iwate University, Morioka, Japan; gakugaku@iwate-u.ac.jp -->
<!--  -- Rebecca Selden, Wellesley College, Wellesley, MA, USA; bselden@wellesley.edu -->
<!--  -- Julia Xue, Shanghai Jiao Tong University; juliaxue@ouc.edu.cn -->

@@

# Abstract
<!-- *An opening sentence that sets the question that you address and is comprehensible to the general reader, background content specific to this study, results, and a concluding sentence. It should be a single paragraph.* -->

Regulatory boundaries and species distributions are often not aligned, especially in the case of marine species crossing multiple Economic Exclusive Zones (EEZs). This represents a challenge for fisheries management as policies tend to focus at the national level, yet collaborations are needed to maximize long-term ecological, social and economic benefits. Here, we combined species distributions and the spatial delimitation of EEZs to identify the number of commercially exploited marine species that are shared between neighboring nations. We found that 67% of the world's identified fished species are transboundary (n = 633) and were responsible for 80% of the global-EEZs catches (48 million tons) and 78.5% of fisheries revenue (USD 65 billion) in the last decade. Some countries received over 90% of their catch and economic benefits from a few shared resources. Our analysis suggests that fisheries targeting transboundary species appear worse off than those mainly catching non-transboundary species. Our study has direct implications for the management of transboundary species, pointing to the urgent need for strengthened international cooperation in the management of these fisheries.

# Introduction

@@

We overlaid the known distribution of 938 commercially valuable marine species responsible for an average of 96.5% of global EEZ catch reconstructed by the *Sea Around Us* at the species level between 2005 and 2014, and 241 EEZs of 198 coastal territories. We define a ‘stock’ unit as a species in an EEZ, instead of using a population ecology definition (genetically or morphological distinct units) due to lack of such biological information for almost all the species included in this study [@Teh:2015gd]. While we acknowledge that a species could have multiple stocks within an EEZ, many fisheries within a country or EEZ are managed at the species instead of stock level. Also, recent research has suggested strong connectivity between stocks, even when separated by thousands of kilometers [@Ramesh:2019va] providing additional ecological ground for our analysis. For this analysis we only considered shared species between neighboring EEZs, rather than the species’ extended distribution (e.g. we did not include the high seas). We rely on multiple data sources including occurrence, distribution models and catch data and only considered a species to be present in a grid cell if all data sources predicted positive occurrence (see Methods).

## Results and Discussion

@@

![**Figure 1.** Number of transboundary species and their contribution to global fisheries catches and revenue. A) The number of species and amount of revenue are represented by color coding of EEZs and land polygons, respectively. B) Contribution of transboundary species to regional revenue (left) and catch (right). Regions according to the United Nations sub-regions. Points = mean +- sd. Revenue in 2010 real USD.](./Figures/Fig1.png)

We group countries according to the United Nations sub-regions and found that transboundary species are particularly economically important for Northern America (mean revenue = 3,901 $\pm$ 4,497 million USD) and Eastern Asia (2,827 $\pm$ 2,667 million USD). The estimated revenues from transboundary species in these two regions are significantly different from other regions in terms of revenue (one-way ANOVA; *DF*(16,158) = 4.948, *p* < 0.001, $\alpha$ = 0.05) (Fig. 1B). China (USD 6,279 million) in E. Asia and the USA (USD 9,605 million) contribute 55% and 82% to the total revenues from the E. Asia and N. America regions, respectively. In addition to China and the USA, Russia (USD 6,361 million), Peru (USD 5,211) and Japan (USD 3,368) are amongst the top five countries with the most fishing revenue generated from transboundary species in the last decade. These five nations are responsible for 40% of the global fisheries revenue from transboundary species.

Peru and Russia, responsible for the two largest fisheries in the world [@FAO:2018KK], lead transboundary catch statistics harvesting over 5.8 million tons annually each on average between 2005 and 2014 (Fig. S1). Peru's catches are mainly composed of Peruvian anchoveta (*Engraulis ringens*), shared with Chile, and account for 79% of the transboundary species production. Recently these two countries have signed an agreement to work towards standardized stock assessments through coordinated management of this species [@UNDP:2016wp]. Similarly, Russia shares the Alaskan pollock (*Theragra chalcogramma*) fishery with both Japan and the USA celebrating a management agreement with the latter over the Berring Sea since 1988 [@NOAA:gX6CtUTk]. Transboundary species also contribute largely to fisheries in E. Asia (one-way ANOVA; *Df*(16,158) = 2.265, *p* = 0.005, $\alpha$ = 0.05). Particularly, China the world's top fish producer [@FAO:2018KK] getting one third (5.1 million tons) of the total catches from transboundary species, followed by Japan (1.8 million tons) and South Korea (1.06 million tons). Differences in transboundary fisheries can also be illustrated in terms of catch-revenue/EEZ $km^2$ (Fig. 2). As an example, the aggregated area of all EEZs of Northern Europe is 3.3 million $Km^2$, the $5^{th}$ smallest of the 17 groups analyzed (Table S1), however, it has the most revenue (USD 246.7 million per $Km^2$) and catch (260.1 tons per $Km^2$) of transboundary species per EEZ area, but only significantly differing from Sub-Saharan African and Southern Europe countries (*p* < 0.05).

```{r Call_Fig_2, eval = F, fig.width=6, fig.height=4, echo=FALSE, fig.pos= "H", fig.cap=""}

![**Figure 2.** Weighted benefits of transboundary species by Km2 and UN sub-region. A) Million 2010 USD. B) Million tons. Point sub-region average bars +- s.d.](./Figures/Fig2.png)

We determined the exploitation category of each species according to the catch trend within each EEZ (Fig. 3). Previous studies demonstrated that catch trends can be used to infer whether a stock is healthy, over-exploited or collapsed [@Kleisner:2011wn]. Following this concept, Category A would represent fisheries that are re-building or under-exploited, category B would be species that are sustainably harvested, and Category C would be species that are over-exploited. We found significant differences in exploitation categories based on whether or not a species was transboundary (MANOVA, *Df*(2,447) = 27.091, *p* < 0.001). Specifically, transboundary species that are only fished by one country are generally less likely to be over-exploited compared to those shared by neighboring countries (Table S2). Significant differences were found for all categories within transboundary species (one-way ANOVA, *Df*(2,447) = 55.01, *p* $\leq$ 0.001, $\alpha$ = 0.05; Tukey's post hoc test *p* < 0.05) but only Category C for discrete species (one-way ANOVA, *Df*(2,380) = 21.86, *p* < 0.001; $\alpha$ = 0.05; Tukey's post hoc test for Categories C-A and C-B *p* < 0.001, and B-A 0.052). This result suggests that the management of discrete species could be better off as the variation between Categories A and B is closer (*p* = 0.052) than for transboundary species (*p* = 0.001).

@@

![Figure 3. Number of transboundary species by catch category. Category A, fishing between 10%-50% of maximum catch level after a drop larger than 10% of the recorded maximum catch; Category B, fishing over 50% of maximum catch level; Category C, fishing less than 50% of maximum catch](./Figures/Fig3.png)

Empirical analysis suggests that in most cases, management of transboundary species will have better outcomes in terms of fish yields when nations cooperate [@Munro:2002uf]. Examples of successful cooperative-management include Norway and Russia over Atlantic cod (*Gadus morhua*) [@Eide:2013dha] and Namibia and South Africa over hake (*Merluccius spp*) [@Sumaila:2003vw]. Lack of collaboration over shared stocks may threaten stock sustainability, reduce the future profitability potential of the fishery, and create conflict between neighboring nations [@Clark:1980js; @Spijkers:2017ij]. Yet, we acknowledge that cooperation or lack thereof can be a complex process [@Jensen:2015cf], and that there are specific cases where cooperation might not be the best strategy [@Munro:2015df]. 
<!-- Europe has the higest rate of transboundary species per $km^2$ (Figure 2) which is consistent with their Common Fisheries Policy to manage the EEZ as a whole. However, despite some cases of successfull fisheries management, excessive quotas, poorly implemented fishing subsidies, and policy-related issues have failed to control fishing effort threating the sustainabillity of most european stocks [@Khalilian:2010bk].  -->

Transboundary fisheries are abundant and of economic importance for some of the top countries with past fisheries-related conflicts including Canada, the USA, Russia and the European Union (EU) [@Spijkers:2019cz]. For example, since 2007, the EU, Norway, Iceland, and the Faroe Islands have been at odds over the size and relative allocation of total allowable catches for Atlantic mackerel (*Scomber scombrus*) due to a climate-driven distribution shift [@Spijkers:2017ij]. Mackerel is a transboundary species that contributed an average of 598.4 thousand tons to catches (8.19% of total) and USD 850.2 million in revenue (7.17% of total) to the countries involved in the dispute between 2005 and 2014. The relatively small contribution of mackerel to the involved countries suggests that even a small fraction of the year's catch can lead to important conflicts between neighboring countries. Climate change is expected to continue changing the distribution and share proportion of fish stocks among countries up to a degree that new transboundary species will emerge [@Pinsky:2018cb], while some species will exit EEZs [@Oremus:2020id]. Exploring the detailed impacts of climate change in share species is key to develop local adaptation methods to minimize the effect of such distributional change on the conservation of species as well as the economy and food security of coastal nations.

Over 130 marine fish species occur in tropical and subtropical waters around the world [@Nelson:2016kk], from highly migratory species associated to pelagic-oceanic ecosystems like tunas (*Thunnus sp.*), to less mobile reef-associated species like greater amberjack (*Seriola dumerili*), to species found in demersal ecosystems like gilthead seabream (*Sparus aurata*). Species associated to pelagic-oceanic ecosystems are the only group whose EEZ range (i.e., the number of EEZs where the species occur as transboundary) was significantly different from other groups (one-way ANOVA; *DF*(5,597) = 53.82, *p* < 0.001, $\alpha$ = 0.05; Tukey's post hoc test *p* < 0.05) with a median of 40 EEZs per species. The median for species of all other ecosystem preferences is close to, or less than, 20, as many of these species have a narrower distribution or are less mobile (Fig. 4).

@@

img <- readPNG(here("./Figures/Fig4.png"))
grid.raster(img)
```

![Figure 4. Number of EEZs shared by transboundary species. A) Number of EEZs by species organized by associated habitat according to FishBase. Showing only species that share > 20 EEZs. B) Catchs of main fishing countries fishing each ecosystem preference's top five species in terms of shared EEZs (shown in bold in A) between 2005 and 2014.](./Figures/Fig4.png)

Due to their large migration patterns and presence in areas beyond national jurisdiction, many highly migratory species are managed by Regional Fisheries Management Organizations (RFMOs). Fifteen major RFMOs in the world oversee the management of 99 species [@CullisSuzuki:2010fi]. We estimated that 81.8% of these species are shared across neighboring RFMOs, with the International Commission for the Conservation of Atlantic Tunas (ICCAT) sharing over 40 species with neighboring RFMOs (Fig. 5). The effectiveness of RFMOs has been previously questioned [@CullisSuzuki:2010fi]. However, despite a global expansion of the industrial fleet in the last decade [@Coulter:2020gr], a recent analysis of 40 tuna and billfish stocks suggests that most stocks are sustainable harvested with only three currently considered as overfished [@Pons:2016ep].

```{r Call_Fig5, eval = F, fig.width=6, fig.height=4, echo=FALSE, fig.pos= "H", fig.cap="Number of shared species by RFMOs. Analysis included with 85 species managed by Fourteen RFMOs"}

img <- readPNG(here("./Figures/Fig5.png"))
grid.raster(img)
```

![Figure 5. Number of shared species across RFMOs. Analysis included 99 species managed by Fourteen RFMOs according to Cullis-Suzuk and Pauly, 2010.](./Figures/Fig5.png)

On the other hand, species with narrower ranges are often managed at the national level due to localized populations [@Payne:2008vm]. However, the potential boundary-crossing of these species must not be ignored. The world of fisheries is highly connected through larvae dispersal and so, even centralized populations could be supplying populations thousands of kilometers away [@Ramesh:2019va]. In addition, coral regions often share multiple species, making coordination among countries key to achieve fisheries and conservation goals [@Payne:2008vm]. Many coastal communities, particularly in developing regions, are highly dependent on subsistence fishing for food security and livelihoods [@CisnerosMontemayor:2016gq; @Wabnitz:2018gf]. For example, all of the top five countries fishing transboundary reef and pelagic-oceanic associated species (Fig. 3B) are among the highest fish consumption countries [@FAO:2018KK]. In these countries, both pelagic and reef fishes contribute substantially to both local food security and livelihood [@Cabral:2018hc]. In the case of Kiribati, the Solomon Islands and Papua New Guinea, they are parties to the Pacific Nauru Agreement that sustainable manage 30% of global skipjack tuna [@Aqorau:2018bh], estimated to produce $2 billion worth of fish every year, representing a role model for transboundary fisheries management [@Hanich:2010ht].

# Conclusions

@@

# Materials and Methods

@@

We developed a four-criteria methodology to determine whether or not a species can be considered transboundary. Only species that meet all criteria were considered as transboundary and the analysis was undertaken only within the EEZs of coastal states. In some cases, criteria can also be interpreted as indices to measure the uncertainty in the analysis as they provide a continuous index (e.g. from 0 to 1) rather than a discrete, yes or no, index.

@@

We estimated the economic contribution in 2010 real USD of transboundary species for each country using global ex-vessel price data [@Tai2017]. The database we draw from includes ex-vessel price from multiple sources and a structured interpolation method (e.g. similar countries, species) to fill in data gaps [@Sumaila:2015uc]. The contribution of transboundary species was derived by aggregating the catch of species occurring within the EEZs of countries sharing that species. For this study, we did not include catch nor revenue from areas beyond national jurisdiction. Finally, we used the catch data to determine the exploitation category of each species within an EEZ. Although this method has previously been used to estimate stock status [@Grainger:1996tk], the categories presented here are intended as catch trends, and not fishing status for each species as many environmental and social-economic factors (e.g. temperature, markets, fishing policies) affect fishing effort and catch volume [@Pauly:2013gs; @Branch:2008kk]. We only assessed species within each EEZ for which at least ten years of data were available between the first and last reported catch and with at least five consecutive years of data. Three final categories were drawn for each species depending on the catch volume within each EEZ (e.g. present, maximum and minimum EEZ's historical catch) and the year (e.g. year of maximum historical catch of the species within that EEZ) (Table 2) [@Kleisner:2011wn]. Following this concept, Category A would represent fisheries that are re-building or under-exploited, Category B would be species that are sustainable harvested, and Category C would be species that are over-exploited. Finally, we reported the predominant category over the last ten years of catch data (2005-2014).

## Statistical analysis

All analyses were run using the statistical software *R version 3.5.2 (2018-12-20)* with the packages *data.table* [@Packagedatatable:2019uh], *janitor* [@PackagejanitorSim:2018ur], *wesanderson* [@Packagewesanderson:2018vn], *rfishbase* [@PackagerfishbaseR:2019th], *R.matlab* [@PackageRmatlabRe:2018un], *sf* [@PackagesfSimpleF:2018vp], *sp* [@PackagespClasses:2019tn], *tidiverse* [@PackagetidyverseE:2017vq], *tidytext* [@PackagetidytextTe:2019vk], and *zoo* [@PackagezooSInfr:2019tm]. All code is available at https://github.com/jepa/FishForVisa. We performed a series of one-way analysis of variance (one-way ANOVA) and Multivariate analysis of variance (MANOVA) to determine statistically significant differences between the means of different groups (e.g., geographical regions, species, exploitation categories) of transboundary species and their contribution to catch and fishing revenue [@Chambers:1992up; @Krzanowski:1990wm]. We used the standard *R* functions for the ANOVA and MANOVA after testing for assumptions. In cases where the ANOVA assumptions were not clearly meet we ran the non-parametric version Kruskal-Wallis Rank Sum Test to confirm the results [@Hollander:2013wz].

@@

# Data Availability
The Sea Around Us datasets used are publicly available at http://www.seaaroundus.org/, the occurrence dataset is all publicly available (see Methods), the ENM-Nereus data can be accessed upon request to G.R. All code is available at https://github.com/jepa/FishForVisa

@@

# References

# Acknowledgments
We thank T. Clark and A. Tidd for helpful comments and discussions. G.R. was supported by the EO Wilson foundation. C.W. was supported by the Walton Family Foundation (grant 2018-1371), the David and Lucile Packard Foundation (grant 2019-68336), and the Gordon and Betty Moore Foundation (grant GBMF5668.02). W.W.L.C was supported by the NSERC Discovery Grant and the SSHRC Partnership.

# Authors contributions
J.P.A. Conceptualization, methodology, formal analysis, writing original draft. G.R. Methodology, data curation, writing - review & editing. C.W. Methodology, writing - review & editing. W.W.L.C. Supervision, writing - review & editing. All authors provided valuable input to the study conceptualization and discussed the results and analysis.

# Additional Information  
**Competing interests**: No declared.

## Supplements

@@

![Figure S1. Number of transboundary species per EEZ and their contributions to countries’ EEZ catch. The number of species is displayed in the EEZ while catch (in thousand tons) is displayed in the land polygon.](./Figures/FigS1.png)

The number of gridcells in which a species is present (*Area Index*) is expected to directly influence whether or not a species is considered transboundary. Therefore, we tested the sensitivity of our results to the *Area Index* by estimating the number of transboundary species along a gradient of values (Fig. S2). As expected, using a more relaxed value (e.g. the species is present within 10% or more of the EEZ) will result in a greater number of transboundary species (median of 40 EEZs shared per species), while a more restrictive value (e.g., 50%) results in a median estimate of 4. The current analysis utilized a 25% *Area Index* threshold (Fig. S2 - darker histogram).
